---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

The data is unziped and loaded with the following script.  The zip file is assumed to be in your working directory.

```{r preprocess}
require(dplyr, quietly = T, warn.conflicts = F)
require(ggplot2, quietly = T)
require(xtable, quietly = T, warn.conflicts = F)

unzip("activity.zip")
data <- read.csv("activity.csv", stringsAsFactors = F)
```

## What is mean total number of steps taken per day?
```{r dailysteps}
stepData <- group_by(data, date)
dailyTotal <- summarize(stepData, n = sum(steps, na.rm = T))
names(dailyTotal) <- c("Date","Steps")
dailyMean <- mean(dailyTotal$Steps)
dailyMedian <- median(dailyTotal$Steps)
```

Total Steps per Day:

```{r results="asis"}
print(xtable(dailyTotal), type="html")
```

Histogram of Total Steps per Day

```{r}
p <- ggplot(data=dailyTotal, aes(dailyTotal$Steps)) + 
  geom_histogram(col="red", 
                 fill="green", 
                 bins = 30,
                 alpha = .2) + 
  labs(title="Total Steps per Day") +
  labs(x="Total Steps", y="Count")
p
```

Mean Steps per Day : `r dailyMean`

Median Steps per Day: `r dailyMedian`


## What is the average daily activity pattern?

```{r daily}
dailyData <- group_by(data, interval)
dailyIntervalTotal <- summarize(dailyData, n = mean(steps, na.rm = T))
names(dailyIntervalTotal) <- c("Interval","Steps")
maxSteps <- which.max(dailyIntervalTotal$Steps)
maxInterval <- dailyIntervalTotal$Interval[maxSteps]
```

Average steps taken per time interval:

```{r}
p <- ggplot(data=dailyIntervalTotal, aes(x=Interval, y=Steps)) + 
  geom_line() + 
  labs(title="Avg Daily Steps") +
  labs(x="Time Interval", y="Avg Steps")
p
```

Maximum time interval for daily average steps: `r maxInterval`

## Imputing missing values

```{r imputing}
missingCount <- sum(is.na(data$steps))

imputeData <- data

for(i in which(is.na(imputeData$steps))) {
     imputeData$steps[i] <- dailyIntervalTotal$Steps[dailyIntervalTotal$Interval == imputeData$interval[i]]
}

imputeStepData <- group_by(imputeData, date)
imputeDailyTotal <- summarize(imputeStepData, n = sum(steps, na.rm = T))
names(imputeDailyTotal) <- c("Date","Steps")
imputeDailyMean <- mean(imputeDailyTotal$Steps)
imputeDailyMedian <- median(imputeDailyTotal$Steps)
options(scipen = 3)

```

Total count of missing values: `r missingCount`

Missing Values imputed using Interval means.

Imputed Histogram of Total Steps per Day

```{r}
p <- ggplot(data=imputeDailyTotal, aes(imputeDailyTotal$Steps)) + 
  geom_histogram(col="red", 
                 fill="green", 
                 bins = 30,
                 alpha = .2) + 
  labs(title="Total Steps per Day") +
  labs(x="Total Steps", y="Count")
p
```

Imputed Mean Steps per Day : `r imputeDailyMean`

Imputed Median Steps per Day: `r imputeDailyMedian`

The imputed values seems to skew things towards the original median, causing a large spike there.

## Are there differences in activity patterns between weekdays and weekends?

```{r weekday}
data <- mutate(data,
     weekday = ifelse(weekdays(as.Date(date,"%Y-%m-%d")) %in% c("Saturday","Sunday"),"weekend","weekday"))

weekday <- group_by(data, interval, weekday)
weekday <- summarize(weekday, n = mean(steps, na.rm = T))
names(weekday) <- c("Interval", "Weekday", "Steps")


p <- ggplot(data=weekday, aes(x=Interval, y=Steps)) + 
  geom_line() + 
  labs(title="Weekday Avg Daily Steps") +
  labs(x="Time Interval", y="Avg Steps") +
  facet_wrap(~Weekday, nrow=2)
p

```
